% !TEX root = ../main.tex
%
\pdfbookmark[0]{Remerciements}{Remerciements}
\addchap*{Remerciements}

The first thing I would like to say is: \textit{Gracias a la vida, que me ha dado tanto}. I am thankful to have closed this cycle of my life, from which I have learned both professionally and personally. I am thankful to have been surrounded by wonderful people, who gave me the energy to move forward. I am thankful to have lived long enough to be here and to be able to look at the sky. I thank you from the bottom of my heart.

I would like to thank my teachers, I feel very fortunate to have worked with you. Pierre, I thank you for all the confidence you have had in my project, my ideas, and my goals. Without your support I could not have come this far, and it is partly thanks to you. Merci beaucoup. Luciana, I thank you for all the love and dedication you have shown me, I will never forget all those details. Without your welcome my Latin heart would be colder, and with you I have felt this warmth that makes me feel at home. Muito obrigado. I would also like to thank Nicolás, because you have been an important person in this journey. Without you, this cycle could not have started, and likewise, it could not have ended. So many conversations we had, which gave me the strength to keep going when I felt stuck. Muchas gracias.

I would like to thank everyone who welcomed me at the Sorbonne Université, because each one contributed to the progress of this journey. So many people who trusted me to start on my teaching path, who gave me the wings to begin a great challenge. Thank you so much.

I would like to thank everyone who has welcomed me in Paris. Although many people come and go, each one will have a part of my heart, and I am grateful for the love they have given me. Without these people, everything would be more difficult.

I would like to thank my family. Alba, quién sabe en qué momento de tu vida estés leyendo esto o probablemente ya te lo habré leído, pero quiero que sepas lo agradecido y orgulloso que estoy de ser tu papá, has sido una hija maravillosa, quién me ha brindado su amor incondicional. Cami, te quiero agradecer por haber emprendido este viaje a mi lado, has sido una de las grandes artífices de esta aventura, la cual ha culminado gracias al cariño y apoyo que me has dado, sin dudas todo sería distinto sin ti. Papá, te quiero agradecer todo el apoyo y confianza que has tenido conmigo, desde que emprendí este viaje siento que hemos pasado un período maravilloso juntos. Sofi, Amandi, Iñaki y Victor, gracias por todo el cariño vertido todas las veces que los visito en nuestra querida patria. Mamá, ya no puedo agradecerte, lamentablemente ya no estás acá, pero te dedico estas palabras de agradecimiento por haberme hecho quién soy. Gracias totales, les amo con todo mi corazón.

Thanks to the Chilean government. This thesis was funded by the National Agency for Research and Development National Agency for Research and Development (ANID) / Scholarship Program / DOCTORADO BECAS CHILE / 2018 - 72190551. I want to thank the project ANID FONDECYT Nº11190314, Chile. This material is based upon work supported by the Google Cloud Research Credits program with the award GCP19980904.